---
title: "Data Import"
author: "Mingyin Wang"
date: 2024-09-18
output: github_document
---

This document contains the code for Homework 1 in P8105

##Problem 1

```{r setup, echo = FALSE, message = FALSE}
library(tidyverse)
```


```{r}
#load the penguins dataset
data("penguins", package = "palmerpenguins")
penguins
```


```{r}
# Code to get the numbers of rows, columns and mean of flipper_length_mm column
dataset_rows <- nrow(penguins)
dataset_cols <- ncol(penguins)
mean_flipper_length = mean(penguins$flipper_length_mm, na.rm = TRUE)
```



```{r yx_scatter}
#Make a scatterplot of flipper_length_mm (y) vs bill_length_mm (x) and save it
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = "species")) + 
  geom_point() + labs(
    x = " Bill length in mm", 
    y = "Flipper length in mm", 
    title = "Scatterplot of Flipper Length vs Bill Length"
  )
ggsave("penguins_scatterplot.pdf", height = 4, width = 6)
```


## Question 2
```{r}
# make a data frame
set.seed(1)  
df <- data.frame(
  random_sample = rnorm(10), 
  logical_vector = rnorm(10) > 0,
  character_vector = sample(letters, 10),
  factor_vector = factor(sample(c("Level1", "Level2", "Level3"), 10,  replace = TRUE))  
)

```

```{r}
# take the mean of each variable
mean_random_sample = mean(df %>% pull(random_sample) )
mean_logical_vector = mean(df %>% pull(logical_vector) )
mean_character_vector = mean(df %>% pull(character_vector) )
mean_factor_vector = mean(df %>% pull(factor_vector) )
```
The random_sample and logical_vector work, since random_sample is numeric and logical_vector contains true and false, which are 1 and 0 respectively. 

```{r, result = 'hide'}
#convert variable to numeric
numeric_logical_vector = as.numeric(df %>% pull(logical_vector))
numeric_character_vector = as.numeric(df %>% pull(character_vector))
numeric_factor_vector = as.numeric(df %>% pull(factor_vector))
```
For the Logical vector, true becomes 1 and false becomes 0, then we could calculate the mean

For the character vector, the conversion does not work and returns NA values, since characters cannot be directly converted to numeric.

For the factor vector,the conversion to numeric works, but the numeric values correspond to the factor levels, level 1, level 2, level 3. Then, we could take the average of factor vector, but it is not meaningful. 







